---
title: "Figs for W2W Year 1 meeting"
author: "Matt Savoca"
date: "2024-10-21"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r cars}

source("Util.R")


```

## Summarizing results from the larger size class 

Samples analyzed at Hopkins Marine Station under microscope

First a plot of the summary data

```{r pressure, echo=FALSE}
# Prep summary data

Opt_micro_beach_cut <- Opt_micro_beach_summ %>% 
  select(Site_ID, Sample_ID, Volume_Pumped_L, Total_Count:Total_Foams) %>% 
  mutate(Sample_type = "beach")

Opt_micro_river_cut <- Opt_micro_river_summ %>% 
  select(Site_ID, Sample_ID, Volume_Pumped_L, Total_Count:Total_Foams) %>% 
  mutate(Sample_type = "river")

Opt_micro_ocean_cut <- Opt_micro_ocean_summ %>% 
  select(Site_ID, Sample_ID, Volume_Pumped_L, Total_Count:Total_Foams) %>% 
  mutate(Sample_type = "ocean")


Opt_micro_all_cut <- rbind(Opt_micro_beach_cut, Opt_micro_river_cut, Opt_micro_ocean_cut) %>% 
  mutate(Site_type = case_when(
    grepl("SRR|PRR", Sample_ID, ignore.case = TRUE) ~ "agricultural",
    grepl("CRR|SLR", Sample_ID, ignore.case = TRUE) ~ "urban",
    TRUE ~ NA_character_  # Use NA_character_ for character NA values
  ),
  Volume_Pumped_L = as.numeric(Volume_Pumped_L),
  Plastics_per_L = Volume_Pumped_L/Total_Count
  )


# Convert the dataframe to long format
long_data <- Opt_micro_all_cut %>% 
  select(-Total_Count) %>%  # Remove Total_Count
  pivot_longer(cols = Total_Fibers:Total_Foams, 
               names_to = "Plastic_Type", 
               values_to = "Value") %>%   # Convert to long format
  mutate(Plastic_Type = case_when(
    Plastic_Type == "Total_Fibers" ~ "fiber",
    Plastic_Type == "Total_Fragments" ~ "fragment",
    Plastic_Type == "Total_Films" ~ "film",
    Plastic_Type == "Total_Nurdles" ~ "nurdle",
    Plastic_Type == "Total_Foams" ~ "foam",
    TRUE ~ Plastic_Type  # This is a catch-all for any other values
  ),
    Sample_type = case_when(
      grepl("Blank", Sample_ID, ignore.case = TRUE) ~ "blank",
      TRUE ~ Sample_type  # Keep original Sample_type if not a blank
    )
  ) %>% 
    filter(!is.na(Value))  # Remove rows where Value is NA

# First, ensure Sample_type is a factor and set the order
long_data <- long_data %>%
  mutate(Sample_type = factor(Sample_type, levels = c("blank", unique(Sample_type[Sample_type != "blank"]))))

# Create a stacked bar plot with the custom palette
Opt_summ_stacked_bar_plot <- ggplot(long_data, aes(x = Sample_type, y = Value, fill = Plastic_Type)) +
  geom_bar(stat = "identity", position = "stack") +  # Stacked bars
  scale_fill_manual(values = custom_palette) +  # Apply custom color palette
  labs(x = "Sample Type", y = "Plastic Count", fill = "Plastic Type") +
  theme_classic(base_size = 18) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

# Display the plot
Opt_summ_stacked_bar_plot


# Calculate proportions
long_data_prop <- long_data %>%
  group_by(Sample_type) %>%
  mutate(Total = sum(Value)) %>%
  ungroup() %>%
  mutate(Proportion = Value / Total)

# Create the plot
Opt_summ_prop_bar_plot <- ggplot(long_data_prop, aes(x = Sample_type, y = Proportion, fill = Plastic_Type)) +
  geom_bar(stat = "identity", position = "stack") +  # Stacked bars
  scale_fill_manual(values = custom_palette) +  # Apply custom color palette
  labs(x = "Sample Type", y = "Proportion of Plastic Types", fill = "Plastic Type") +
  theme_classic(base_size = 18) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels
  scale_y_continuous(labels = scales::percent_format())  # Format y-axis as percentages

# Display the plot
Opt_summ_prop_bar_plot



Opt_summ_AgvsUrban <- long_data %>%
  filter(!is.na(Site_type)) %>%  # Remove rows where Site_type is NA
  ggplot(aes(x = Site_type, y = Value, fill = Plastic_Type)) +
  geom_bar(stat = "identity", position = "stack") +  # Stacked bars
  scale_fill_manual(values = custom_palette) +  # Apply custom color palette
  labs(x = "Site Type", y = "Plastic count in river water", fill = "Plastic Type") +
  theme_classic(base_size = 18) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

# Display the plot
Opt_summ_AgvsUrban


```

## Summarizing results from the larger size class 

Samples analyzed at Hopkins Marine Station under microscope

Next plots of the particle details from optical microscopy

```{r}

Opt_micro_river_dets_cut <- Opt_micro_river_dets %>% 
  select(-Site_ID, -Original_ID)


Opt_micro_dets_all_cut <- rbind(Opt_micro_river_dets_cut, Opt_micro_ocean_dets, Opt_micro_beach_dets) %>% 
  mutate(Morphology = case_when(
    Morphology %in% c("Fiber", "Fiber Clump", "fiber", "fiber clump") ~ "fiber",
    Morphology %in% c("Fragment", "fragment") ~ "fragment",
    tolower(Morphology) == "film" ~ "film",
    Morphology == "Foams" ~ "foam",
    TRUE ~ as.character(Morphology)  # Keep any other categories as they are
  ),
    Color = case_when(
      Color %in% c("Navy blue", "Blue", "Light Blue", "Navy Blue", "Teal", "blue", "Light blue") ~ "blue",
      tolower(Color) == "clear" ~ "clear",
      Color %in% c("Black", "Gray") ~ "black",
      Color %in% c("Red", "Maroon", "red") ~ "red",
      Color %in% c("Burgendy", "Tan") ~ "brown",
      Color %in% c("White", "White/Blue", "Silver", "white") ~ "white",
      Color %in% c("Green", "Dark Green") ~ "green",
      TRUE ~ as.character(Color)  # Keep any other colors as they are
    )
  )
  


ggplot(Opt_micro_dets_all_cut, aes(x = Approximate_Size_Microns)) +
  geom_density(aes(color = Morphology, fill = Morphology), alpha = 0.5) +
  geom_density(color = "black", size = 1) +  # Add this line for the overall distribution
  geom_rug(aes(color = Morphology), alpha = 0.5) +
  scale_x_log10() +  # Use log scale for x-axis due to potential large range in sizes
  scale_fill_manual(values = custom_palette) +  # Apply custom color palette
  scale_color_manual(values = custom_palette) +  # Apply custom color palette
  labs(x = "Approximate Size (Microns)", 
       y = "Density",
       title = "Distribution of Microplastic Sizes by Morphology") +
  theme_minimal(base_size = 16) +
  theme(legend.position = "bottom")



# Create the stacked bar plot
Bar_color_plot <- ggplot(Opt_micro_dets_all_cut, aes(x = fct_infreq(Color), fill = Color)) +
  geom_bar(position = "stack", color = "black") +  # Use geom_bar to create a bar plot
  scale_fill_manual(values = custom_color_palette) +  # Apply the custom color palette
  labs(x = "Color", y = "Count", fill = "Color") +  # Label the axes
  theme_classic(base_size = 16)  # Apply a classic theme
Bar_color_plot

```

Lab blanks plot

```{r}

```




## Summarizing results from the smaller size class 

Samples analyzed at Moore Institute for Plastic Pollution Research

First a plot of the summary data
```{r}

# Stacked bar plot of count by sample_num, colored by material_class

# First, reorder sample_num based on the total count per sample
combined_data_part_summ$sample_num <- fct_reorder(combined_data_part_summ$sample_num, combined_data_part_summ$count, .fun = sum, .desc = TRUE)

stacked_bar_material_class_plot <- ggplot(combined_data_part_summ, aes(x = sample_num, y = count, fill = material_simple)) +
  geom_bar(stat = "identity") +  # 'stat="identity"' is used because we are plotting precomputed counts
  labs(x = "Sample Number", y = "Count", fill = "Material Class") +  # Label the axes and legend
  theme_classic(base_size = 14)  # Apply a classic theme

# Display the plot
stacked_bar_material_class_plot


```
```{r}
library(dplyr)
library(knitr)

# Filter for rows where material_simple is 'plastic'
plastic_data <- combined_data_part_summ[combined_data_part_summ$material_simple == "plastic", ]

# Calculate total plastic count per sample
plastic_summary_by_sample <- plastic_data %>%
  group_by(sample_num) %>%
  summarise(
    total_plastic_count = sum(count)
  )

# Calculate overall total plastic count
total_plastic_overall <- sum(plastic_data$count)

# Add a row for overall total at the bottom
overall_summary <- data.frame(
  sample_num = "Overall",
  total_plastic_count = total_plastic_overall
)

# Combine the sample summary and the overall summary
plastic_summary <- bind_rows(plastic_summary_by_sample, overall_summary)

# Print the table using knitr::kable for a nice format
plastic_summary %>%
  kable(
    col.names = c("Sample Number", "Total Plastic Count"),
    caption = "Summary of Total Plastic Count by Sample",
    align = "c"
  )


```
Now plastic deets for the 54-500u range

```{r}


# Filter for rows where material_simple is "plastic"
plastic_data <- combined_data_part_summ %>%
  filter(material_simple == "plastic")

# Summarize the count of plastic by material_class
plastic_class_summary <- plastic_data %>%
  group_by(material_class) %>%
  summarise(total_plastic_count = sum(count)) %>%
  arrange(desc(total_plastic_count))  # Arrange by count from high to low

plastic_histogram <- ggplot(plastic_class_summary, aes(x = reorder(material_class, -total_plastic_count), y = total_plastic_count, fill = material_class)) +
  geom_bar(stat = "identity", color = "black") +
  labs(x = "Plastic type", y = "Total Count") +
  theme_classic(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),  # Adjust vjust to move labels closer
        legend.position = "none")  # Remove the legend



# Display the plot
plastic_histogram

ggsave("plastic_histogram.pdf", width = 18, height = 8)

```


```{r}
library(dplyr)
library(knitr)
library(kableExtra)

table_output <- combined_data_part_summ %>%
  filter(material_simple == "plastic") %>%
  group_by(material_class) %>%
  summarize(total_count = sum(count)) %>%
  ungroup() %>%
  mutate(proportion = total_count / sum(total_count)) %>%
  arrange(desc(total_count)) %>%
  mutate(proportion = scales::percent(proportion, accuracy = 0.1))

# Calculate the total sum
total_sum <- sum(table_output$total_count)
total_proportion <- scales::percent(1, accuracy = 0.1)

# Add the total row
table_output <- bind_rows(
  table_output,
  tibble(
    material_class = "Total",
    total_count = total_sum,
    proportion = total_proportion
  )
)

# Create a formatted table
kable(table_output, 
      col.names = c("Material Class", "Total Count", "Proportion"),
      align = c("l", "r", "r")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
  row_spec(0, bold = TRUE) %>%
  row_spec(nrow(table_output), bold = TRUE) %>%
  add_header_above(c(" " = 1, "Plastic Types" = 2))


```

Details on small plastic shape

```{r}
ggplot(combined_data_part_dets, aes(x = min_length_um, y = max_length_um, color = material_simple)) +
  geom_point(alpha = 0.5) +  # Set alpha to 0.5 for 50% transparency
  scale_x_log10() +  # Use log scale for x-axis due to potential large range
  scale_y_log10() +  # Use log scale for y-axis due to potential large range
  labs(x = "Minimum Length (µm)", 
       y = "Maximum Length (µm)", 
       color = "Material Type",
       title = "Marticle Max Length vs Min Length") +
  theme_minimal() +
  theme(legend.position = "bottom")




ggplot(combined_data_part_dets, aes(x = area_um2, y = perimeter_um, color = material_simple)) +
  geom_point(alpha = 0.5) +  # Set alpha to 0.5 for 50% transparency
  scale_x_log10() +  # Use log scale for x-axis due to potential large range
  scale_y_log10() +  # Use log scale for y-axis due to potential large range
  labs(x = "Area (µm^2)", 
       y = "Perimeter (µm)", 
       color = "Material Type",
       title = "Particle area vs perimeter") +
  theme_minimal() +
  theme(legend.position = "right")

```
Aspect ratio plot
```{r}

# Filter the data for only plastic rows
filtered_data <- combined_data_part_dets %>%
  filter(material_simple == "plastic")

# Calculate the median aspect_ratio for plastic
median_aspect_ratio <- median(filtered_data$aspect_ratio, na.rm = TRUE)

# Create the density plot with vertical dashed line at the median
aspect_ratio_density_plot <- ggplot(filtered_data, aes(x = aspect_ratio, fill = material_simple, color = material_simple)) +
  geom_density(alpha = 0.4) +  # Add density plot with transparency
  geom_rug(aes(color = material_simple), sides = "b", size = 0.8) +  # Add a rug plot below
  geom_vline(aes(xintercept = median_aspect_ratio), linetype = "dashed", color = "black", size = 1) +  # Add dashed vertical line at median
  xlim(1, 10) +  # Limit the x-axis to 1 to 10
  labs(x = "Aspect Ratio", y = "Density", title = "Density Plot of Microplastic Aspect Ratio") + 
  theme_classic(base_size = 14)  # Apply a classic theme

# Display the plot
aspect_ratio_density_plot



# Calculate the median aspect_ratio for plastic
median_circularity <- median(filtered_data$circularity, na.rm = TRUE)

# Create the density plot with vertical dashed line at the median
circularity_density_plot <- ggplot(filtered_data, aes(x = circularity, fill = material_simple, color = material_simple)) +
  geom_density(alpha = 0.4) +  # Add density plot with transparency
  geom_rug(aes(color = material_simple), sides = "b", size = 0.8) +  # Add a rug plot below
  geom_vline(aes(xintercept = median_circularity), linetype = "dashed", color = "black", size = 1) +  # Add dashed vertical line at median
  #xlim(1, 10) +  # Limit the x-axis to 1 to 10
  labs(x = "Circularity", y = "Density", title = "Density Plot of Microplastic Circularity") + 
  theme_classic(base_size = 14)  # Apply a classic theme

# Display the plot
circularity_density_plot

```






